\showyourwork: a workflow for open source scientific articles

### 2021-12-23 src/ms.tex
intro

@@

Introduction coming soon.

>>>>

\usepackage{enumitem}
\setlist[enumerate]{wide=0pt,widest=99,leftmargin=\parindent,labelsep=*}
\setlist[itemize]{wide=0pt,widest=99,leftmargin=\parindent,labelsep=*}

@@

This paper introduces \showyourwork, a workflow that enables the creation and distribution of fully reproducible and open source scientific articles. 
As astronomical software becomes increasingly more complex, and as research results become increasingly more interdependent, it is crucial to ensure the validity and correctness of papers published in the field.
However, the current peer review system is simply not set up to do this, since checking all of the results in a paper would require the painstaking and methodical review of all of the paper's methods—which usually means scrutinizing all of the code used to generate the figures, tables, and other quantities in the paper. 
In practice, this is virtually impossible for three reasons:

\begin{enumerate}
    %
    \item Modern codebases can be very large and often require deep familiarity with the software to use—not to mention review them.
    %
    \item  Writing a paper in astronomy is rarely ever done in a linear, procedural fashion: the codebase is constantly changing, and the state of the code when (say) Figure 1 was produced may be very different from that when (say) Figure 2 was made. 
    Moreover, many results depend on the execution of lengthy pipelines with intermediate steps, each potentially requiring manual tinkering that is not always documented and may be difficult to exactly replicate.
    %
    \item The majority of astronomical code is not open source and simply cannot be vetted by third parties. 
    While there has been a marked increase in the number of open source astronomical tools in recent years (e.g., astropy, exoplanet, emcee, exofast...), most code associated with the generation of the results in individual papers is not open source; readers are often expected to take it on faith that there are no bugs in that code, or that the code works exactly as described in the text, with no pitfalls or missing details. 
    Even when the code is made publicly available, e.g., by being published on \texttt{GitHub}, it is often not documented sufficiently to enable one to execute it and reproduce the paper's results out-of-the-box. 
    And even with proper documentation, the code may require external dependencies, custom virtual environments, or access to closed-source datasets that make it difficult or impossible for a third party to replicate it.
    %
\end{enumerate}

\showyourwork was designed to tackle these three issues, making it easy to develop, publish, and distribute truly open and reproducible research papers in astronomy. 
It exists as a \texttt{GitHub} template repository, which can be cloned at the click of a button to set up a new article. 
Users then add their LaTeX manuscript, bibliography, scripts used in the generation of the paper's figures, an anaconda environment specification, and instructions on how to download any external datasets required by the figures. 
Every time the user pushes a new commit to \texttt{GitHub}, the article is automatically built on the cloud using \texttt{GitHub Actions} and the resulting PDF is pushed to a separate branch of the repository. 
The build step—which sets up the conda environment, generates all figures from scratch (with intelligent caching), and compiles the PDF—acts as a unit test for the paper. 
If it passes, the paper is (by definition) reproducible.

The workflow works out of the box for simple projects, in which each figure is generated by running a given \texttt{Python} script. 
But it also works for more complicated pipelines, such as projects that depend on many intermediate steps or those that require running expensive simulations on clusters. 
The workflow interfaces directly with \texttt{Zenodo}, allowing users to automatically upload the results of simulations so that expensive build steps can be bypassed on the cloud. 
In fact, most of the stuff under the hood is there to make the workflow as flexible and customizable as possible; see the documentation for a list of examples of custom workflows.

Papers that use this workflow can be reproduced by cloning the repository and running \texttt{make}. 
Furthermore, these papers include clickable icons next to each of their figures linking to (1) the exact version of the script used to generate them and (2) the exact version(s) of the \texttt{Zenodo}-hosted dataset(s) used in their creation.
